### 2020-07-03 chapters/background.tex
Small corrections

In Section \ref{sec:stream-superset}, we introduced batch and stream processing as two large scale data processing techniques. We defined each and elaborated on the fundamental differences between both paradigms. We conclude that there is no best technique and that each serves a different use case.

>>>>

In Section \ref{sec:stream-superset}, we introduced batch and stream processing as two large scale data processing techniques. We defined each, elaborated on the fundamental differences between both paradigms but also introduced the superset relation there is between stream and batch processing.

### 2020-07-03 chapters/conclusions.tex
Small corrections

    \item experiments on real data that did not show reasonable results but where we provide insights and possible hypothesis to test for future work
\end{enumerate}

@@

In this thesis we had to deal with the multiple test problem and apply a multiple test correction. In our case, we used Holm-Bonferroni due to its simplicity and because it was similar to past work \cite{SAMM}. There are many other multiple test correction methods, and we leave the testing of each as future work.

>>>>

    \item experiments on real data that showed unexpected results but where we provide insights and possible hypothesis to test for future work
\end{enumerate}

@@

In this thesis we had to deal with the multiple test problem and apply a multiple test correction. In our case, we used Holm-Bonferroni due to its simplicity and because it controls the family wise error rate with a larger power than some other methods (thus providing a good trade-off between simplicity and power). There are many other multiple test correction methods, and we leave the testing of each as future work.

### 2020-07-03 chapters/introduction.tex
Small corrections

This section merely sums up how each of the following chapters contributes to this thesis.

In Chapter \ref{chap:background} of this document, we brief the reader with the necessary background and relevant concepts needed to understand the following work, namely the topics of Large Scale Data Processing, Sliding Window Aggregations and Outlier Detection, as well as all the sub-topics inherent to them.

@@

In Chapter \ref{chap:my-work}, we discuss initial solutions and reasons to discard them before presenting our distribution monitoring method.

In Chapter \ref{chap:validation}, we describe the set of experiments made and the datasets used. We start with single feature experiments on synthetic data, then multi feature analysis and finally move on to multi feature analysis on real data.

In Chapter \ref{chap:conclusion}, we draw our final conclusions. We reiterate our hypothesis and research questions and provide answers to them. Finally, we conclude by summarizing our contributions and make explicit current issues to be addressed as future work.

>>>>

This section outlines the content of each chapter of this thesis.

In Chapter \ref{chap:background} of this document, we brief the reader with the necessary background and relevant concepts needed to understand the following work, namely the topics of Large Scale Data Processing, Sliding Window Aggregations and Outlier Detection, as well as all relevant sub-topics inherent to them.

@@

In Chapter \ref{chap:my-work}, we discuss initial approaches and reasons to discard them before presenting our proposed solution.

In Chapter \ref{chap:validation}, we describe the datasets used and the set of experiments that were performed to validate our method. We start with single feature experiments on synthetic data, then multi feature analysis and finally move on to multi feature analysis on real data.

In Chapter \ref{chap:conclusion}, we draw our final conclusions. We reiterate our hypothesis and research questions and provide answers to them. Finally, we conclude by summarizing our contributions and, make explicit current issues to be addressed as future work.

### 2020-07-03 chapters/sota.tex
Small corrections

@@

%\subsubsection{Automatic Model Monitoring for Data Streams}
%ALSO a 2 WINDOW MODEL
%\textcolor{red}{TODO: write this}

@@

\textcolor{red}{TODO Find and add versions of Sliding Bloom Filter and Sliding CMS}

@@

In Section \ref{sec:outliers}, we reviewed outlier detection algorithms. However, we found most of them inapplicable in our scenario, because they were resource intensive (both time and memory wise), provided no measure of how divergent and outlier was or because they lost track of their reference period.

@@

We then discussed probabilistic data structures (Section \ref{sec:pds}) and their sliding window implementations (Section \ref{sec:sliding-pds}). Despite being efficient, both time and memory wise, each probabilistic data structure allows for one type of aggregation, and none of them happened to be of particular use for us.

>>>>

Track changes is off
Everyone
Track changes for everyone
You
Track changes for You
Guests
Track changes for guests
pedro.cardoso: How is this error calculated?
Jun 18, 2020 2:52 PM
You: Same as above
Jun 20, 2020 5:09 PM • Edit • Delete
Hit Enter to reply
pedro.cardoso: This 3rd option is under-explained compared to the other 2. Please elaborate
Jun 18, 2020 2:53 PM
You: Not sure myself, that is why is so poorly written. Im going to revisit this
Jun 19, 2020 4:42 PM • Edit • Delete
Hit Enter to reply
pedro.cardoso: Thought: This reads like leftover from the original work. Is there a way to better relate your investigation into SWAGs into the methodology you ended up taking?
Jun 30, 2020 10:19 AM
pedro.cardoso: Please revisit this entire section. Particularly the applicability sections of each algorithm in the context of your current approach.
Jun 30, 2020 10:20 AM
Hit Enter to reply
pedro.cardoso: Remove these TODOs. If you want keep a note of them elsewhere, but not as text render being rendered.
Jul 3, 2020 5:28 PM
pedro.cardoso: You will forget to do it with other tasks still needed to be done.
Jul 3, 2020 5:29 PM
Hit Enter to reply
pedro.cardoso: "because they evolved their reference period"
Jul 3, 2020 5:30 PM
Hit Enter to reply
pedro.cardoso: Where is Model Monitoring?
Jul 3, 2020 5:31 PM
Hit Enter to reply
Current file
Overview
16

@@

%\subsubsection{Automatic Model Monitoring for Data Streams}
%ALSO a 2 WINDOW MODEL

@@

@@

In Section \ref{sec:outliers}, we reviewed outlier detection algorithms. However, we found most of them inapplicable in our scenario, because they were resource intensive (both time and memory wise), provided no measure of how divergent and outlier was or because they evolved their reference period.

@@

We then discussed probabilistic data structures (Section \ref{sec:pds}) and their sliding window implementations (Section \ref{sec:sliding-pds}). Despite being efficient, both time and memory wise, each probabilistic data structure allows for one type of aggregation, and none of them happened to be of particular use to us.

### 2020-07-03 chapters/validation.tex
Small corrections

Feature x1 has less alerts in this experiment where we use equal reference and target periods. However, the alarms last for about two months. This is unexpected as we predicted we would see no alarms or few ones with short duration. Likewise, we observe that for feature x2 (Figure \ref{fig:merchant2-x2-correctedpvalues}) and x3 (Figure \ref{fig:merchant2-x3-correctedpvalues}) we still have many alarms and some lasting up to one month.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x1-correctedpvalues.pdf}
      \caption{Merchant feature x1 corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x1-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x2-correctedpvalues.pdf}
      \caption{Merchant feature x2 corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x2-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x3-correctedpvalues.pdf}
      \caption{Merchant feature x3 corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x3-correctedpvalues}
    \end{center}
\end{figure}
We do not have a clear explanation for this and due to time constraints we leave further tests to future work. We suspect it may be related to the way EMAs work: they keep every past event in the aggregation state. In other words, despite the fact EMAs apply a smoothing factor to the aggregation state for each new event, the previous events never have their contribution set to zero, even though the weights tend to zero.

@@

\section{Conclusions}
In this section, we summarize our experimental tests, results and found issues.

@@

Finally, we moved on to real data. Due to time constraints and the scope of this thesis, we experimented only with data from one merchant. In our first experiment with real data, we ran our method with what we thought was an ideal set of parameters and produced a lot of alarms. A close analysis of the reference and target time-series of each feature revealed that some alarms indeed made sense, as this dataset contained a lot of drift. In our last experiment, we attempted to reduce concept drift to a maximum by using the same period as reference and target. To our surprise, we still produced a large number of long lasting alarms. We defer the investigation of the root cause of the problem to future work and leave a few untested hypothesis as solutions for this issue, like the possibility that despite applying a smoothing factor to the aggregation state for each new event, EMAs never forget past events. In other words, the previous events never have their contribution set to zero, even though the weights tend to zero. Additionally, our sampling may have just been insufficient. To test this, an experiment with lower $\gamma$ and hence higher number of samples should be carried out.

>>>>

Feature x1 has less alerts in this experiment where we use equal reference and target periods. However, the alarms last for about two months. This is unexpected as we predicted we would see no alarms or few ones with short duration. Likewise, we observe that for feature x2 (Figure \ref{fig:merchant2-x2-correctedpvalues}) and x3 (Figure \ref{fig:merchant2-x3-correctedpvalues}) we still have many alarms and some lasting up to one month.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x1-correctedpvalues.pdf}
      \caption{Merchant feature x1 corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x1-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x2-correctedpvalues.pdf}
      \caption{Merchant feature x2 corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x2-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant2-x3-correctedpvalues.pdf}
      \caption{Merchant feature x3 corrected p-values with equal reference and target periods}
      \label{fig:merchant2-x3-correctedpvalues}
    \end{center}
\end{figure}
We do not have a clear explanation for this and due to time constraints we leave further tests to future work. We suspect the issue may be that there is a difference between the batch and streaming phases. In batch we have subsets of events and compute EMA-like histogram aggregations based on those. In streaming, however, we have a continuous stream of data and never expire the 6\% of data that in batch is not present, because it is a finite window.

@@

\section{Conclusions}
In this section, we summarize our experimental tests, results and uncovered issues.

@@

Finally, we moved on to real data. Due to time constraints and the scope of this thesis, we experimented only with data from one merchant. In our first experiment with real data, we ran our method with what we thought was an ideal set of parameters and produced a lot of alarms. A close analysis of the reference and target time-series of each feature revealed that some alarms indeed made sense, as this dataset contained a lot of drift. In our last experiment, we attempted to reduce concept drift to a maximum by using the same period as reference and target. To our surprise, we still produced a large number of long lasting alarms. We defer the investigation of the root cause of the problem to future work and leave a few untested hypothesis as solutions for this issue, like the possibility that the EMA-like histograms in batch work with subsets of data and the streaming ones work with infinite amounts of data. In other words, for the streaming EMAs, the previous events never have their contribution set to zero, even though the weights tend to zero, whereas the batch EMAs worked with a finite dataset. Additionally, our sampling may have been insufficient. To test this, an experiment with lower $\gamma$ and hence higher number of samples should be carried out.
